Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-14 paper/paper.tex
pass on technique

\Mar{Lucas/Guilherme, verifique se isto esta correto}
Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}. A
Suricata rule is divided in three
parts---action, header, and rule options~\cite{suri-rule-format}. The
action part appears as the first word in the rule description. In this
case, \CodeIn{alert}. An action denotes the task that needs to be
executed if the rule pattern is satisfied. In this case, a message
will be sent to sys admins if the rule pattern is satisfied. Other
options of actions include pass, drop, and reject. The
header comes after the action in the rule description. It restricts
the information flow covered by the rule. For this rule, the header is
\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which instructs Suricata
to inspect \CodeIn{tcp} traffic flowing from any port in the
home network to any other address outside the home network. The
variables \CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are configurable. The rule
options come after the header. It is a semi-colon-separated sequence
of key-value pairs characterizing the attack pattern. For example,
\Mar{Lucas/Guilherme (IMP): explique o que cada propriedade faz...}

@@

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established.

Aware of the
protocol, an attacker sends multiple SYN packets to different ports of
a server, often using fake IP addresses. Without proper protection,
the server accepts the connection requests and eventually legitimate
connection requests cannot be satisfied due to resource exhaustion.

@@

>>>>

\Mar{Lucas/Guilherme, verifique se isto esta correto}
Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}. A
Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The action part appears as the first
word in the rule description. In this case, \CodeIn{alert}. An action
denotes the task that needs to be executed if the rule pattern is
satisfied. In this case, a message will be sent to sys admins if the
rule pattern is satisfied. Other options of actions include pass,
drop, and reject. The header comes after the action in the rule
description. It restricts the information flow covered by the
rule. For this rule, the header is \CodeIn{tcp \$HOME\_NET any ->
  \$EXTERNAL\_NET any}, which instructs Suricata to inspect
\CodeIn{tcp} traffic flowing from any port in the home network to any
other address outside the home network. The variables
\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
configurable. The rule options come after the header. It is a
semi-colon-separated sequence of properties\Mar{should we call it an
  option or property?}, where each property is a key-value pair
describing a distinct characteristic of the attack.

@@

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with a SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established. Aware of the protocol, an attacker
sends multiple SYN packets to different ports of a server, often using
fake IP addresses. Without proper protection, the server accepts the
connection requests and eventually legitimate connection requests
cannot be satisfied due to resource exhaustion.

Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
this attack. The critical parts are the option \CodeIn{flags: S,12},
which identifies a SYN packet and the option \CodeIn{threshold: type
  both, track by\_dst, count 5000, seconds 5}, indicating that a high
volume of such packets should be requested in a short period of time.

@@

\tname{} is a search-based algorithm that uses positive and negative
examples to synthesize signature-based rules, such as those from Snort
and Suricata. \tname{} synthesizes rules in two steps. First, it
produces an overspecified rule to capture the negative example
provided as input. Then, it uses positive examples to eliminate
unecessary options from the rule. The goal is to create a rule that
captures only the negative traffic.

\subsection{Problem Characterization}

\subsection{Fitness Functions}

\subsection{Search}
